---
title: "analysis"
format: html
---

```{r}

# turn off scientific notation
options(scipen = 999)

# load Packages
source("fxns.R")
library(tidyverse)
library(ellmer)
library(glue)

movies <- read.csv("data/raw/horror_movie_locations.csv", header = FALSE) |> 
  clean_names() |> 
  rename(
    title = v1,
    state = v2,
    rating = v3,
    multiple_locations = v4,
    description = v5
  ) |> 
  # some state names are mispelled
  mutate(
    state = trimws(state),
    state = case_when(
      state == "Flordia" ~ "Florida",
      state == "District of Colombia" ~ "District of Columbia",
      state == "West Virgina" ~ "West Virginia",
      state == "New Mexco" ~ "New Mexico",
      state == "Illlinois" ~ "Illinois",
      state == "Lousiana" ~ "Louisiana",
      state == "Oho" ~ "Ohio",
      state == "Kanasa" ~ "Kansas",
      TRUE ~ state
    )
  ) |> 
  mutate(
    year = as.numeric(str_extract(title, "\\d{4}")),
    title = str_remove(title, " \\(\\d{4}\\)$"),
    multiple_locations = case_when(
      multiple_locations == "Multiple Locations" ~ TRUE,
      TRUE ~ FALSE
    ),
    word_count = str_count(description, boundary("word")),
    year_bin = case_when(
      year >= 2020 ~ "over_20",
      year >= 2010 & year < 2020 ~ "x10_20",
      year >= 2000 & year < 2010 ~ "x00_10",
      year >= 1990 & year < 2000 ~ "x90_00",
      year >= 1980 & year < 1990 ~ "x80_90",
      year >= 1970 & year < 1980 ~ "x70_80",
      year >= 1960 & year < 1970 ~ "x60_70",
      year >= 1950 & year < 1960 ~ "x50_60",
      year >= 1940 & year < 1950 ~ "x40_50",
      year >= 1930 & year < 1940 ~ "x30_40",
      year < 1930 ~ "under_30"
    ),
    ai_info = paste0('The movie is "', title,'" filmed in ',year,'. This is the IMDb summary: ', description)
  )

# movies where the year wasn't listed (none in MD)
no_year_movies <- movies |> filter(is.na(year)) |> pull(title)

# movies filmed in maryland
movies_maryland <- movies |> 
  filter(state == "Maryland")

```


## Non AI analysis

```{r}

movies_by_state <- movies |> 
  group_by(state) |> 
  summarize(
    count = n(),
    avg_rating = mean(rating, na.rm = TRUE),
    med_rating = median(rating, na.rm = TRUE),
  )

movies_by_state_year <- movies |> 
  group_by(state, year_bin) |> 
  summarize(
    count = n(),
    avg_rating = mean(rating, na.rm = TRUE),
    med_rating = median(rating, na.rm = TRUE),
  )
```

## Defining structured output

```{r}

type_characteristics <- type_object(
  primary_sub_genre = type_enum(
    c(
      "phobia",
      "madness",
      "home invasion",
      "slasher",
      "backwoods horror",
      "torture",
      "body horror",
      "splatter",
      "cannibal",
      "extreme",
      "zombies",
      "virus",
      "vampire",
      "werewolf",
      "classic and mythological",
      "neo-monsters",
      "nature",
      "giant creatures",
      "small creatures",
      "sci-fi and aliens",
      "ghost and spirits",
      "haunted house",
      "possession",
      "devils, demons, and hell",
      "witches",
      "supernatural power",
      "horror comedy",
      "parody horror",
      "Lovecraftian/Cosmic horror",
      "gothic horror",
      "found footage",
      "folk horror",
      "post-apocalyptic"
    ),
    "Identify the primary sub-genre of the movie"
  ),
  secondary_sub_genre = type_enum(
    c(
      "phobia",
      "madness",
      "home invasion",
      "slasher",
      "backwoods horror",
      "torture",
      "body horror",
      "splatter",
      "cannibal",
      "extreme",
      "zombies",
      "virus",
      "vampire",
      "werewolf",
      "classic and mythological",
      "neo-monsters",
      "nature",
      "giant creatures",
      "small creatures",
      "sci-fi and aliens",
      "ghost and spirits",
      "haunted house",
      "possession",
      "devils, demons, and hell",
      "witches",
      "supernatural power",
      "horror comedy",
      "parody horror",
      "Lovecraftian/Cosmic horror",
      "gothic horror",
      "found footage",
      "folk horror",
      "post-apocalyptic"
    ),
    "Identify the secondary sub-genre of the movie. You may not choose {primary_sub_genre} as it has been chosen already."
  ),
  tertiary_sub_genre = type_enum(
    c(
      "phobia",
      "madness",
      "home invasion",
      "slasher",
      "backwoods horror",
      "torture",
      "body horror",
      "splatter",
      "cannibal",
      "extreme",
      "zombies",
      "virus",
      "vampire",
      "werewolf",
      "classic and mythological",
      "neo-monsters",
      "nature",
      "giant creatures",
      "small creatures",
      "sci-fi and aliens",
      "ghost and spirits",
      "haunted house",
      "possession",
      "devils, demons, and hell",
      "witches",
      "supernatural power",
      "horror comedy",
      "parody horror",
      "Lovecraftian/Cosmic horror",
      "gothic horror",
      "found footage",
      "folk horror",
      "post-apocalyptic"
    ),
    "Identify the tertiary sub-genre of the movie. You may not choose {primary_sub_genre} or {secondary_sub_genre} as they have been chosen already."
  ),
  protagonist_type = type_enum(
    c(
      "human",
      "human with extra powers",
      "possessed human",
      "non-human entity",
      "unknown"
    ),
    "The protagonist is a "
  ),
  antagonist_type = type_enum(
    c(
      "human",
      "human with extra powers",
      "possessed human",
      "non-human entity",
      "unknown"
    ),
    "The antagonist is a "
  ),
  protagonist_gender = type_enum(
    c("male", "female", "unknown"),
    "The protagonist's gender is "
  ),
  antagonist_gender = type_enum(
    c("male", "female", "androgynous", "unknown"),
    "The antagonist's gender is "
  ),
  protagonist_age = type_enum(
    c("under 18", "18-30", "30-50", "50+", "unknown"),
    "The protagonist is within the following age range"
  ),
  antagonist_age = type_enum(
    c("under 18", "18-30", "30-50", "50+", "unknown"),
    "The antagonist is within the following age range"
  ),
  setting_time = type_enum(
    c("past", "present", "future"),
    "The movie was filmed in {year}. The movie is set in the "
  ),
  setting_location = type_enum(
    c(
      "house",
      "forest",
      "big city",
      "small town",
      "rural",
      "outer space",
      "unknown"
    ),
    "The movie is primarily located in the "
  )
)

```


## Declare global chat objects

```{r}

system_prompt_movie <- readLines("prompt.md")

# GPT models
if (Sys.getenv("OPENAI_API_KEY") == "") {
  stop("ERROR: OPENAI_API_KEY environment variable is not set")
} else {
  gpt_api <- Sys.getenv("OPENAI_API_KEY")
}

chat_gpt_41 <- chat_openai(
    system_prompt = system_prompt_movie,
    model = "gpt-4.1",
    api_key = gpt_api
  )

chat_gpt_41_mini <- chat_openai(
    system_prompt = system_prompt_movie,
    model = "gpt-4.1-mini",
    api_key = gpt_api
  )

chat_gpt_5_mini <- chat_openai(
    system_prompt = system_prompt_movie,
    model = "gpt-5-mini",
    api_key = gpt_api
  )

chat_gpt_5 <- chat_openai(
    system_prompt = system_prompt_movie,
    model = "gpt-5",
    api_key = gpt_api
  )

# claude models

if (Sys.getenv("ANTHROPIC_API_KEY") == "") {
  stop("ERROR: ANTHROPIC_API_KEY environment variable is not set")
} else {
  anthropic_api <- Sys.getenv("ANTHROPIC_API_KEY")
}

chat_claude_sonnet_37 <- chat_anthropic(
    system_prompt = system_prompt_movie,
    model = "claude-3-7-sonnet-20250219",
    api_key = anthropic_api
  )

chat_claude_sonnet_4 <- chat_anthropic(
    system_prompt = system_prompt_movie,
    model = "claude-sonnet-4-20250514",
    api_key = anthropic_api
  )

# Gemini models

if (Sys.getenv("GOOGLE_API_KEY") == "") {
  stop("ERROR: GOOGLE_API_KEY environment variable is not set")
} else {
  gemini_api <- Sys.getenv("GOOGLE_API_KEY")
}

chat_gemini_flash_preview <- chat_google_gemini(
    system_prompt = system_prompt_movie,
    mode = "gemini-2.5-flash-preview-09-2025",
    api_key = gemini_api,
  )

chat_gemini_pro <- chat_google_gemini(
    system_prompt = system_prompt_movie,
    mode = "gemini-2.5-pro",
    api_key = gemini_api
  )

chat_gemini_flash <- chat_google_gemini(
    system_prompt = system_prompt_movie,
    mode = "gemini-2.5-flash",
    api_key = gemini_api
  )

valid_models <- list("gpt-4.1" = chat_gpt_41, 
                      "gpt-4.1-mini" = chat_gpt_41_mini,
                      "gpt-5-mini" = chat_gpt_5_mini,
                      "gpt-5" = chat_gpt_5,
                      "claude-3-7-sonnet-20250219" = chat_claude_sonnet_37,
                      "claude-sonnet-4-20250514" = chat_claude_sonnet_4,
                      "gemini-2.5-pro" = chat_gemini_pro,
                      "gemini-2.5-flash" = chat_gemini_flash,
                      "gemini-2.5-flash-preview-09-2025" = chat_gemini_flash_preview)

```


## Ellmer function

```{r}

# Functions
address_ellmer <- function(chat, ai_info) {

  # Wrap entire function in tryCatch to prevent stopping execution
  tryCatch({

    # Input validation
    if (!chat %in% names(valid_models)) {
      stop(paste("ERROR: chat must be one of:", toString(names(valid_models))))
    }

    if (missing(ai_info) || is.null(ai_info) || ai_info == "") {
      stop("ERROR: ai_info must be provided and cannot be empty")
    }
    chat_obj <- valid_models[[chat]]

    # Execute chat with error handling and retry logic
    result <- NULL
    max_retries <- 3
    retry_count <- 0
    last_error <- NULL

    while (retry_count < max_retries && is.null(result)) {
      tryCatch({
        chat_obj$set_turns(list())        
        result <- chat_obj$chat_structured(ai_info, type = type_characteristics, echo = "all")

        # Validate result
        if (is.null(result) || (is.character(result) && result == "")) {
          warning("Empty result received from ", chat, " API")
          result <- NULL
          last_error <- "Empty result received from API"
        }

      }, error = function(e) {
        retry_count <<- retry_count + 1
        last_error <<- e$message

        if (retry_count < max_retries) {
          warning("Attempt ", retry_count, " failed for ", chat, " API: ", e$message, ". Retrying...")
          Sys.sleep(2^retry_count)  # Exponential backoff
        }
      })
    }

    # Final validation
    if (is.null(result)) {
      return(paste("ERROR: Failed to get valid response from", chat, "API after", max_retries, "attempts. Last error:", last_error))
    }

    return(result)

  }, error = function(e) {
    # Catch-all for any unexpected errors
    return(paste("ERROR: Unexpected error in address_ellmer:", e$message))
  })
}

```


## Test different models

```{r}

movie_test <-
  movies |> 
  # Popular movie that I know well, not popular movie from MD, popular movie from MD
  filter(title %in% c("Smile","Camp Killer","The Blair Witch Project")) |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x)),
    cost_gpt_41 = chat_gpt_41$get_cost(include = "all"),
    gpt_41_mini = map(ai_info, ~ address_ellmer("gpt-4.1-mini", .x)),
    cost_gpt_41_mini = chat_gpt_41_mini$get_cost(include = "all"),
    gpt_5_mini = map(ai_info, ~ address_ellmer("gpt-5-mini", .x)),
    cost_gpt_5_mini = chat_gpt_5_mini$get_cost(include = "all"),
    gpt_5 = map(ai_info, ~ address_ellmer("gpt-5", .x)),
    cost_gpt_5 = chat_gpt_5$get_cost(include = "all"),
    sonnet_37 = map(ai_info, ~ address_ellmer("claude-3-7-sonnet-20250219", .x)),
    cost_claude_sonnet_37 = chat_claude_sonnet_37$get_cost(include = "all"),
    sonnet_4 = map(ai_info, ~ address_ellmer("claude-sonnet-4-20250514", .x)),
    cost_claude_sonnet_4 = chat_claude_sonnet_4$get_cost(include = "all"),
    gemini_pro = map(ai_info, ~ address_ellmer("gemini-2.5-pro", .x)),
    cost_gemini_pro = chat_gemini_pro$get_cost(include = "all"),
    gemini_flash = map(ai_info, ~ address_ellmer("gemini-2.5-flash", .x)),
    cost_gemini_flash = chat_gemini_flash$get_cost(include = "all"),
    gemini_flash_preview = map(ai_info, ~ address_ellmer("gemini-2.5-flash-preview-09-2025", .x)),
    cost_gemini_flash_preview = chat_gemini_flash_preview$get_cost(include = "all")
  ) 

#  unnest_wider(movie_gemini, names_sep = "_")

```

Claude doesn't support my type_object and GPT-5 is too new

Smile:
GPT 4.1 - 11/11 subgenres were madness, possession, phobia
GPT 4.1 mini - 11/11 subgenres were madness, possession, phobia
Gemini Pro - 10/11 subgenres were madness, possession, phobia, wrong setting
Gemini Flash - 10/11 subgenres were madness, devils.., phobia, wrong setting
Gemini Flash preview - 10/11 subgrenres were madness, possession, ghosts, wrong setting

Camp Killer:
GPT 4.1 - 10/11 subgenres were slasher, madness, phobia of which slasher and madness were right. Should've included comedy or satire
GPT 4.1 mini - 10/11 subgenres were slasher, home invasion, madness of which slasher was right. Should've included comedy or satire
Gemini Pro - 9/11 subgenres were slasher, madness, phobia, got antagonist age and gender wrong
Gemini Flash - 7/11 subgenres were slasher, madness, phobia, missed protagonist and antagonist age and gender
Gemini Flash preview - 9/11 subgenres were slasher, madness, phobia, missed protagonist and antagonist gender

The Blair Witch Project:
GPT 4.1 - 11/11 subgenres were found footage, folk horror, phobia
GPT 4.1 mini - 11/11 subgenres were found footage, folk horror, phobia
Gemini Pro - 11/11 subgenres were found footage, witches, folk horror
Gemini Flash - 11/11 subgenres were found footage, witches, folk horror
Gemini Flash preview - 11/11 subgenres were found footage, witches, folk horror

```{r}

movie_maryland_test <- movies_maryland |> 
  sample_n(10) |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x)),
    cost_gpt_41 = chat_gpt_41$get_cost(include = "all"),
    gpt_41_mini = map(ai_info, ~ address_ellmer("gpt-4.1-mini", .x)),
    cost_gpt_41_mini = chat_gpt_41_mini$get_cost(include = "all")
  )

```

Last Night at Terrace Lanes:
GPT 4.1 - 11/11
GPT 4.1 mini - 9/11
Multiple Maniacs:
GPT 4.1 - 11/11
GPT 4.1 mini - 10/11
The Scooby Doo Project
GPT 4.1 - 11/11
GPT 4.1 mini - 9/11



## Maryland analysis

```{r}

movies_maryland_classified <- movies_maryland |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

test <- movies_maryland_classified |> 
  unnest_wider(gpt_41, names_sep = "_")

protagonist_gender <- test |> 
  group_by(gpt_41_protagonist_gender) |> 
  summarize(
    n = n()
  )

antagonist_gender <- test |> 
  group_by(gpt_41_antagonist_gender) |> 
  summarize(
    n = n()
  )

protagonist_age <- test |> 
  group_by(gpt_41_protagonist_age) |> 
  summarize(
    n = n()
  )

genres <- test |> 
  pivot_longer(cols = c(gpt_41_primary_sub_genre, gpt_41_secondary_sub_genre, gpt_41_tertiary_sub_genre), names_to = "column", values_to = "genre") |> 
  count(genre, sort = TRUE)

top_genres <- genres |> 
  slice(1:10) |> 
  pull(genre)

genres_over_time <- test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  pivot_longer(cols = c(gpt_41_primary_sub_genre, gpt_41_secondary_sub_genre, gpt_41_tertiary_sub_genre), names_to = "column", values_to = "genre") |> 
  count(year_bin,genre) |> 
  filter(genre %in% top_genres)

setting_over_time <- test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  count(year_bin,gpt_41_setting_location)

gender_over_time <- test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  count(year_bin,gpt_41_protagonist_gender)

p <- ggplot(gender_over_time, aes(x = year_bin, y = n, color = gpt_41_protagonist_gender)) +
  geom_line(size = 1) +
  geom_point(size = 1.5)

ggplotly(p)

```

```{r}

movie_states <- movies |> 
  pull(state) |> 
  unique()

# manual batching

safe_address_ellmer <- safely(address_ellmer, otherwise = NULL)

movies_new_mexico_classified <- movies |> 
  filter(state == "New Mexico") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_oregon_classified <- movies |> 
  filter(state == "Oregon") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_maine_classified <- movies |> 
  filter(state == "Maine") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_arizona_classified <- movies |> 
  filter(state == "Arizona") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_washington_classified <- movies |> 
  filter(state == "Washington") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_georgia_classified <- movies |> 
  filter(state == "Georgia") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_ohio_classified <- movies |> 
  filter(state == "Ohio") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_nevada_classified <- movies |> 
  filter(state == "Nevada") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_illinois_classified <- movies |> 
  filter(state == "Illinois") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_massachusetts_classified <- movies |> 
  filter(state == "Massachusetts") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_jersey_classified <- movies |> 
  filter(state == "New Jersey") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_pennsylvania_classified <- movies |> 
  filter(state == "Pennsylvania") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_louisiana_classified <- movies |> 
  filter(state == "Louisiana") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_florida_classified <- movies |> 
  filter(state == "Florida") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_new_york_classified <- movies |> 
  filter(state == "New York") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_texas_classified <- movies |> 
  filter(state == "Texas") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_california_classified <- movies |> 
  filter(state == "California") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movies_michigan_classified <- movies |> 
  filter(state == "Michigan") |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

lesser_states <- movies_by_state |> 
  arrange(-count) |> 
  slice(20:54) |> 
  pull(state)

movies_north_carolina_virgin_islands <- movies |> 
  filter(state %in% lesser_states) |> 
  mutate(
    gpt_41 = map(ai_info, ~ address_ellmer("gpt-4.1", .x))
  )

movie_test <- bind_rows(
  movies_maryland_classified,
  movies_new_mexico_classified,
  movies_north_carolina_virgin_islands,
  movies_michigan_classified,
  movies_ohio_classified,
  movies_georgia_classified,
  movies_washington_classified,
  movies_arizona_classified,
  movies_maine_classified,
  movies_oregon_classified,
  movies_nevada_classified,
  movies_illinois_classified,
  movies_massachusetts_classified,
  movies_jersey_classified,
  movies_pennsylvania_classified,
  movies_louisiana_classified,
  movies_florida_classified,
  movies_new_york_classified,
  movies_texas_classified,
  movies_california_classified
) |> 
  unnest_wider(gpt_41, names_sep = "_")

# genres overall

movie_test_genres <- movie_test |> 
  pivot_longer(cols = c(gpt_41_primary_sub_genre, gpt_41_secondary_sub_genre, gpt_41_tertiary_sub_genre), names_to = "column", values_to = "genre") |> 
  filter(genre == "found footage")

top_genres <- movie_test_genres |> 
  slice(1:10) |> 
  pull(genre)

ne_states <- c("Virginia","West Virginia","District of Columbia","Delaware","Maryland","New Jersey","Pennsylvania","New York","Connecticut","Massachusetts","Rhode Island","New Hampshire","Vermont","Maine")

dmv <- c("Maryland","District of Columbia","Virginia")

movie_test_genres_over_time <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  pivot_longer(cols = c(gpt_41_primary_sub_genre, gpt_41_secondary_sub_genre, gpt_41_tertiary_sub_genre), names_to = "column", values_to = "genre") |> 
  filter(year %in% 1970:2020) |> 
  count(genre, year_bin) |> 
  group_by(year_bin) |> 
  mutate(
    rank = dense_rank(desc(n)),
    total = sum(n)/3,
    prop = 100*n/total
  ) |> 
  ungroup() |> 
  filter(genre %in% c("found footage","backwoods horror","phobia","witches","folk horror")) |> 
  select(year_bin, genre, prop) |> 
  pivot_wider(
    names_from = genre,
    values_from = prop
  ) |> 
  arrange(year_bin)

p <- ggplot(movie_test_genres_over_time, aes(x = year_bin, y = rank, color = genre)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) 

ggplotly(p)

```

## Maryland comparison

Maryland has 39% of movies w female protagonist whereas median is 36.8% of all time
12.5% of its movies in the forest whereas 9.5% is the median
5.95% of its movies are found footage compared to 1.4% is the median. Only states ahead of its raw count (10) are NY and Cali
All 10 of these films were released after The Blair Witch Project

```{r}

movie_gender_by_state <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  filter(year %in% 1970:2020) |> 
  count(state, gpt_41_protagonist_gender) |> 
  group_by(state) |> 
  mutate(
    prop = n/sum(n),
    med = median(prop)
  ) |> 
  ungroup() |> 
  filter(gpt_41_protagonist_gender == "female")

movie_setting_by_state <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  count(state, gpt_41_setting_location) |> 
  group_by(state) |> 
  mutate(
    prop = n/sum(n),
  ) |> 
  ungroup() |> 
  filter(gpt_41_setting_location == "forest")

movie_genre_by_state <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  pivot_longer(cols = c(gpt_41_primary_sub_genre, gpt_41_secondary_sub_genre, gpt_41_tertiary_sub_genre), names_to = "column", values_to = "genre") |> 
  #filter(year %in% 1900:2020) |> 
  count(state, genre) |> 
  group_by(state) |> 
  mutate(
    total = sum(n)/3,
    prop = round(100*n/total,2)
  ) |> 
  ungroup() |> 
  filter(genre %in% c("found footage")) |> 
  arrange(-n)

maryland_found_footage <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  pivot_longer(cols = c(gpt_41_primary_sub_genre, gpt_41_secondary_sub_genre, gpt_41_tertiary_sub_genre), names_to = "column", values_to = "genre") |> 
  filter(year %in% 1970:2020,
         state == "Maryland",
         genre == "found footage") |> 
  pull(title)

maryland_gender_trend <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  count(state, year_bin, gpt_41_protagonist_gender) |> 
  group_by(state, year_bin) |> 
  filter(year_bin %in% 2000:2020) |> 
  mutate(
    prop = n/sum(n),
  ) |> 
  ungroup() |> 
  filter(gpt_41_protagonist_gender == "female") |> 
  group_by(year_bin) |> 
  mutate(
    med_prop = median(prop)
  ) |> 
  ungroup() |> 
  filter(state == "Maryland")

maryland_gender_trend <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  filter(year_bin %in% 2000:2020) |> 
  count(state, gpt_41_protagonist_gender) |> 
  group_by(state) |> 
  mutate(
    prop = n/sum(n),
    med_prop = median(prop)
  ) |> 
  ungroup() |> 
  filter(gpt_41_protagonist_gender == "female")

movie_setting_trend <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  filter(year_bin %in% 2000:2020) |> 
  count(state, gpt_41_setting_location) |> 
  group_by(state) |> 
  mutate(
    prop = n/sum(n),
    med_prop = median(prop)
  ) |> 
  ungroup() |> 
  filter(gpt_41_setting_location == "forest")
  
p <- ggplot(maryland_gender_trend, aes(x = year_bin)) +
  geom_line(aes(y = prop, color = "Maryland"), size = 1.2) +
  geom_point(aes(y = prop, color = "Maryland"), size = 2) +
  geom_line(aes(y = med_prop, color = "National Median"), linetype = "dashed", size = 1) 

ggplotly(p)

maryland_films <- movie_test |> 
  filter(state == "Maryland") |> 
  arrange(-rating) |> 
  #slice(1:15) |> 
  mutate(
    title = paste0(title," (",year,")"),
    found_footage = case_when(
      gpt_41_primary_sub_genre == "found footage" | gpt_41_secondary_sub_genre == "found footage" | gpt_41_tertiary_sub_genre == "found footage" ~ "Y",
      TRUE ~ "N"
    )
  ) |> 
  select(title, rating, found_footage)
  
```


## DMV genres

Slasher movies were more prevalent in the DMV following SOTL

```{r}

movie_test_genres_dmv <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  pivot_longer(cols = c(gpt_41_primary_sub_genre, gpt_41_secondary_sub_genre, gpt_41_tertiary_sub_genre), names_to = "column", values_to = "genre") |> 
  filter(year %in% 1970:2020,
         state %in% dmv) |> 
  count(genre, year_bin) |> 
  group_by(year_bin) |> 
  mutate(
    rank = dense_rank(desc(n)),
    prop = n/sum(n)
  ) |> 
  ungroup() |> 
  filter(genre %in% c("cannibal","slasher"))

p <- ggplot(movie_test_genres_dmv, aes(x = year_bin, y = prop, color = genre)) +
  geom_line(size = 1) +
  geom_point(size = 1.5)

ggplotly(p)

```


## Setting

Doubled by prevalence from 3.4% of horror movies to 7.2%.

Forest is persistently the 5th most popular by rank

```{r}

setting_over_time <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  count(year_bin,gpt_41_setting_location) |> 
  rename(setting = gpt_41_setting_location) |> 
  filter(year_bin %in% c(1960:2020)) |> 
  group_by(year_bin) |> 
  mutate(
    rank = dense_rank(desc(n)),
    prop = n/sum(n)
  ) |> 
  ungroup()

p <- ggplot(setting_over_time, aes(x = year_bin, y = prop, color = setting)) +
  geom_line(size = 1) +
  geom_point(size = 1.5)

ggplotly(p)

```

## Gender

Female protagonists remained on the rise following 1990, surpassing 50% in the 2020s

```{r}

gender_over_time_maryland <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  rename(gender = gpt_41_protagonist_gender) |> 
  filter(year_bin %in% c(1960:2020),
         state == "Maryland") |> 
  count(year_bin,gender) |> 
  group_by(year_bin) |> 
  mutate(
    rank = dense_rank(desc(n)),
    prop = n/sum(n)
  ) |> 
  ungroup()

gender_over_time <- movie_test |> 
  mutate(
    year_bin = year - (year%%10)
  ) |> 
  count(year_bin,gpt_41_protagonist_gender) |> 
  rename(gender = gpt_41_protagonist_gender) |> 
  filter(year_bin %in% c(1960:2020)) |> 
  group_by(year_bin) |> 
  mutate(
    rank = dense_rank(desc(n)),
    prop = n/sum(n)
  ) |> 
  ungroup()

p <- ggplot(gender_over_time_maryland, aes(x = year_bin, y = prop, color = gender)) +
  geom_line(size = 1) +
  geom_point(size = 1.5)

ggplotly(p)

```

## DMV gender

```{r}

movie_test_genres_dmv <- movie_test |> 
  mutate(
    year_bin = year - (year%%10),
    diff_gender = case_when(
      gpt_41_antagonist_gender == "male" & gpt_41_protagonist_gender == "female" ~ "Y",
    TRUE ~ "N"
    )
  ) |> 
  filter(year %in% 1970:2020,
         state %in% dmv) |> 
  group_by(year_bin) |> 
  summarise(
    total = n(),
    diff_gender = sum(diff_gender == "Y", na.rm = TRUE),
    prop_diff_gender = diff_gender / total
  ) |> 
  ungroup()

p <- ggplot(movie_test_genres_dmv, aes(x = year_bin, y = prop_diff_gender)) +
  geom_line(size = 1) +
  geom_point(size = 1.5)

ggplotly(p)

```

